% Emacs, this is -*-latex-*-
Recent results in coupled or temporal graphical models offer schemes for estimating the relationship structure 
between features when the data come from
related (but distinct) longitudinal sources. A novel application of these ideas is for analyzing group-level differences, i.e., in identifying if {\em trends} of estimated objects (e.g., 
covariance or precision matrices) are different across disparate conditions (e.g., gender or disease). Often, poor effect sizes make detecting the \textit{differential} signal 
over the {\em full} set of features difficult: for example, 
dependencies between only a {\em subset of features} may manifest differently across groups.
In this work, we first give a parametric model 
for estimating trends in the space of $\SPD$ matrices as a function of one or more covariates. We then generalize scan statistics to graph structures, 
to search over distinct subsets of features (graph partitions) whose temporal dependency structure may show statistically 
significant group-wise differences.
We theoretically analyze the Family Wise Error Rate (FWER) and bounds on Type 1 and Type 2 error. 
Evaluating on US census data, we identify groups of states with cultural and legal overlap related to baby name trends and drug usage.
On a cohort of individuals with risk factors for Alzheimer's disease (but otherwise cognitively healthy), 
we find scientifically interesting 
group differences where the default analysis, i.e., models estimated on the {\em full graph}, do not survive reasonable 
significance thresholds. 
