In many real world situations, first-order information in the data is often valuable in identifying group differences. Restricting 
our analysis to only the second-order interactions, i.e., covariances, may be inefficient (or sub-optimal) when the mean signal difference between groups is large. Our construction easily extends to these cases. Particularly, the \textit{product space} over both means and covariances is in $\mathbb{R}^{p} \times \SPD{(p)}$. 
\begin{remark}
The typical GLM on the first order information is defined in the standard Euclidean space. So, computing the regression in the product space 
$\mathbb{R}^{p} \times \SPD{(p)}$ 
amounts to simply 
computing the regression on the first and second order statistics (mean and covariance) separately.
\end{remark}
The above statement suggests that by applying the manifold regression to the covariances and the standard regression model for the means, 
we are directly solving the product space regression problem, incorporating both first and second order statistics.
However, in these cases, the statistic defined above in \eqref{eq:traj} does {\em not} directly take into account the potential difference in means. 
However, given our Normal noise assumption we can easily invoke the standard Gaussian multivariate likelihood statistic for group differences.
\begin{definition}
Let $\hat{\mu}_t, \hat{\Sigma}_t$ be the estimated mean and covariance from the standard linear model and our manifold-covariance GLM respectively. Then the Gaussian likelihood of our data $X$ is
\begin{align}
P(X|\hat{\mu},\hat{\Sigma}) = \prod_{t=1}^{T}\prod_{i=1}^{n_t} P(X_t | N(\hat{\mu}_t,\hat{\Sigma}_t)),
\end{align}
where $X_t$ is the subset of our data collected at timepoint $t$. Additionally, we can define a standard likelihood ratio test statistic as:
\begin{align}
L_{prod} = \frac{P(X_1|\hat{\mu}_1,\hat{\Sigma}_1)P(X_2|\hat{\mu}_2,\hat{\Sigma}_2)}{P(X_{1,2}|\hat{\mu}_{1,2},\hat{\Sigma}_{1,2})}
\end{align}
\end{definition}
This statistic is again $\chi^2_{p}$-distributed \cite{seber2003linear}, and an $\alpha$-level hypothesis test for group difference analysis 
can be defined in the same way as above. While our manifold regression modeling is focused on the case of centered data (where the mean signal may not be 
significantly different between the groups), we use the product space construction, wherever appropriate, in experimental evaluations.
