\chapter{Background}\label{chap:bknd}
Here we will briefly describe some background concepts and ideas that will aid and facilitate discussion of subset identification in later chapters. 
Following some general notation,
we will begin with an overview
of classical probability and statistics,
and a focus on the hypothesis testing
schemes built upon in the third chapter.
This is followed by the basics of Riemannian differential geometry
and the geometry of tensor objects,
the key objects that enable the 
tests and efficiencies described in both 
Chapters \ref{chap:covtraj} and \ref{chap:ott}.
Next we will provide
an overview of optimization tools
and methods particularly suited
and developed for neural network methods,
alongside typical forms and objectives 
in the machine learning literature,
used in Chapter \ref{chap:lcodec}.
We conclude this chapter
with a discussion on optimal transport
methods in deep learning, important
for our discussion of novel optimization schemes
in Chapter \ref{chap:demd}.

\section{General Notations}
The following notations will be standard throughout the sequel, with any particular chapter overloads or redefinitions explicitly mentioned.
\begin{itemize}
\item Calligraphic capital letters such as $\cD, \cM, \cX$ will typically represent spaces.

\item Lowercase letters such as $i,x,n$ will represent vectors as well as scalars (indices, sizes, and dimensions), and subsequently a vector $v \in \RR^d$ would be a $d$-dimensional vector over the reals.

\item Capital letters such as $X,H,T$ will refer to matrices and higher order tensors, as well as sets (including random variables) based on context.

\item We use $\RR^d$ to represent $d$-dimensional vector space over the reals.

\item For learning settings, the training dataset will typically be defined by $S$, with samples $x_i \in S$ in the general setting, and $(x_i,y_i)\in S$ in the supervised setting. Other datasets $T,U$ may also be used and defined in context.

\item Subscripts will denote indexing into sets or vectors, or estimation (in the case of integrals and expectations) over the specified subscript. Superscripts will be used as additional decorators when needed.

\end{itemize}

\section{Probability and Independence}
\input{2_bknd/probstat.tex}

\section{Differential Geometry}\label{sec:diffgeom}
\input{2_bknd/dffgeom.tex}

\subsection{Tensors and their Geometry}
\input{2_bknd/tensors.tex}

\section{Deep Networks, Optimization, and Objectives}
\input{2_bknd/dnnopt.tex}

Each chapter in the sequel
is defined by a unique intersection
of the above ideas, leading to new insights
with respect to the subset selection problem of interest.
The relevant background will be referenced and refreshed 
as needed to ease narrative continuity.